Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-26 NEUROC/sec4-experiments.tex
修改了一些小错误和调整了summary的位置

\paragraph{Summary of Training Time Breakdown}
%
The GNN training was suitable for GPUs.
%
\emph{The edge calculation stage was the main performance bottleneck in most cases}, except for training GNNs with high vertex calculation complexity on low-average-degree graphs.
%
The performance bottleneck in the edge calculation stage depended on the time complexity of the messaging function $\phi$.
%
If the time complexity of $\phi$ was {high}, {$\phi$} dominated the training time of the edge calculation stage.
%
Optimizations should focus on improving its efficiency.
%
Otherwise, the {collection step} and the {aggregation step} dominated the training time.
%
The collection step suffered from lots of data movement.
%
The aggregation step suffered from data synchronization and non-regular data access.

\subsubsection{Performance Bottlenecks in Inference}
%
We also did all the above experiments for Inference.
%
We found that the performance characteristics of GNN inference were highly similar to training on the \emph{layer} level and the \emph{step} level.
%
Here, we used GCN as an example. 
%
We made similar observations for all GNNs. 
%
\figurename~\ref{fig:inference_analysis_on_layer_edge_level_of_gcn} shows the results of GCN inference on the \emph{layer} level and the \emph{step} level.
%
Compared to the training time breakdown of GCN \figurename~\ref{fig:exp_vertex_edge_cal_proportion}(a), 
\figurename~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}(a), and \figurename~\ref{fig:exp_edge_calc_decomposition}(a),
we could conclude that the time breakdowns were very similar.

% GCN Inference结果对比图
\begin{figure}[H]
    \centering
    \subfloat[\emph{layer} level]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[the effect of average degree]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[\emph{edge} level]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_edge_calc_decomposition_gcn.pdf}}
    \caption{Time breakdowns on the layer level and edge level, and the effect of average degree on the edge/vertex calculation of GCN Inference.}
    \label{fig:inference_analysis_on_layer_edge_level_of_gcn}
\end{figure}

@@

The main differences between training and inference were reflected in two aspects: the wall-clock time and the top time-consuming basic operators.
%
The inference time was much less than the training time.
%
Figure~\ref{fig:compare_wall_clock_time_of_training_and_inference} in the response (Figure X in the revised manuscript) compares the wall-clock time of training and inference on the \texttt{amp}, \texttt{amc}, and \texttt{fli} datasets.
%
The results on the other datasets were similar.
%
Since the inference only conducted the forward propagation from the input layer to the prediction layer, the time of inference was very close to the time of the forward phase in training.
%
The inference time was 34\% (GCN), 32\% (GGNN), 25\% (GAT), and 32\% (GaAN) of the training time, averaged over all datasets.

@@

@@

\paragraph{Summary of Memory Usage in Training}
%
The \emph{high} memory expansion ratio severely restricted the data scalability of GNN training.
%
The memory usage mainly came from the intermediate results of the \emph{edge calculation stage}.
%
Fixing the number of vertices, the memory usage increased \emph{linearly} along with the number of edges.
%
Fixing the GNN structure and the hyper-parameters, increasing the dimension of input feature vectors could reduce the memory expansion ratio.
%
To reduce the memory usage of GNN training, optimizations should focus on reducing memory footprints of the edge calculation stage.

@@

@@

\paragraph{Summary of Sampling Techniques in Training}

The sampled subgraphs had lower average degrees than the original graph.
%
With small batch sizes, the sampling techniques could significantly reduce the training time per batch and the peak memory usage.
%
However, small batch sizes could not make full use of the computing power of a GPU.
%
With big batch sizes, the current implementation of the sampling techniques in PyG was inefficient.
%
The time spent on the sampling phase and the data transferring phase even exceeded the training phase.

@@

>>>>

%
We found that the performance characteristics of GNN inference were highly similar to training on the \emph{layer} level and the \emph{step} level.
%
Here, we used GCN as an example. 
%
\figurename~\ref{fig:inference_analysis_on_layer_edge_level_of_gcn} shows the results of GCN inference on the \emph{layer} level and the \emph{step} level.
%
By cross-comparing \figurename~\ref{fig:exp_vertex_edge_cal_proportion}(a), 
\figurename~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}(a), and \figurename~\ref{fig:exp_edge_calc_decomposition}(a),
we could conclude that the time breakdowns were very similar.
%
For the other GNNs, we observed similar phenomena.
%

% GCN Inference结果对比图
\begin{figure}[H]
    \centering
    \subfloat[Layer Level]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[Effects of the average degree]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[Edge Level]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_edge_calc_decomposition_gcn.pdf}}
    \caption{Time breakdowns on the layer level and edge level, and effects of average degree on the edge/vertex calculation of GCN inference.}
    \label{fig:inference_analysis_on_layer_edge_level_of_gcn}
\end{figure}

@@

The main differences between training and inference were reflected in two aspects: the wall-clock time and the top time-consuming basic operators.
%
The inference time was much less than the training time.
%
Figure~\ref{fig:compare_wall_clock_time_of_training_and_inference} compares the wall-clock time of training and inference on the \texttt{amp}, \texttt{amc}, and \texttt{fli} datasets.
%
The results on the other datasets were similar.
%
Since the inference only conducted the forward propagation from the input layer to the prediction layer, the time of inference was very close to the time of the forward phase in training.
%
The inference time was 34\% (GCN), 32\% (GGNN), 25\% (GAT), and 32\% (GaAN) of the training time, averaged over all datasets.

@@

\paragraph{Summary}
% TODO:
The GNN training/inference was suitable for GPUs.
%
\emph{The edge calculation stage was the main performance bottleneck in most cases}, except for training GNNs with high vertex calculation complexity on low-average-degree graphs.
%
The performance bottleneck in the edge calculation stage depended on the time complexity of the messaging function $\phi$.
%
If the time complexity of $\phi$ was {high}, {$\phi$} dominated the training time of the edge calculation stage.
%
Optimizations should focus on improving its efficiency.
%
Otherwise, the {collection step} and the {aggregation step} dominated the training time.
%
The collection step suffered from lots of data movement.
%
The aggregation step suffered from data synchronization and non-regular data access.

@@

@@

\paragraph{Summary of Memory Usage in Training}
%TODO: 
The \emph{high} memory expansion ratio severely restricted the data scalability of GNN training.
%
The memory usage mainly came from the intermediate results of the \emph{edge calculation stage}.
%
Fixing the number of vertices, the memory usage increased \emph{linearly} along with the number of edges.
%
Fixing the GNN structure and the hyper-parameters, increasing the dimension of input feature vectors could reduce the memory expansion ratio.
%
To reduce the memory usage of GNN training, optimizations should focus on reducing memory footprints of the edge calculation stage.

@@

@@

\paragraph{Summary of Sampling Techniques}
% TODO:
The sampled subgraphs had lower average degrees than the original graph.
%
With small batch sizes, the sampling techniques could significantly reduce the training time per batch and the peak memory usage.
%
However, small batch sizes could not make full use of the computing power of a GPU.
%
With big batch sizes, the current implementation of the sampling techniques in PyG was inefficient.
%
The time spent on the sampling phase and the data transferring phase even exceeded the training phase.
